---
title: "Terrorist Ideology, Violence Type, and Selectivity"
author: "Emma Lucas"
description: "A report to identify a causal relationship between terrorist ideology and types of violence as well as selectivity of attacks."
output:
  distill::distill_article:
    self_contained: false
---

## Introduction

For my final project, I was interested in exploring data related to domestic terrorism, violence, and ideology. How does the ideological orientation of domestic terrorist perpetrators influence, if at all, the type of political violence and the selectivity of said violence? In this study, I examined whether or not certain types and selectivity of terrorist violence can indicate patterns and trends within specific ideologies. I hypothesise that less extreme ideologies will have less organised and less wide-scale types of violence and that they will be less selective in their attacks. Less extreme ideologies, while still passionate about their causes, are less likely to take greater steps to ensure their causes. The more the ideology deviates from the norm, the more inclined terrorist organisations are to act out to bring attention to their cause. I also predict that less extreme ideologies will be smaller-scaled in terms of selectivity whereas more extreme ideologies will be larger-scaled and more selective. Neonazis, for instance, will have a set of targeted groups they wish to insue violence on. My explanatory variable of interest is what ideology the terrorist group aligns with. 

## Data Section

The dataset currently holds information on all terrorist attacks from 1965 to 2005 carried out by domestic terrorist organizations in European nations. Each data point represents an individually killed as a directly result of the terrorist attack. To gather the data, researchers relied on pre-existing datasets of terrorist killings (mainly TWEED and Global Terrorism Database), as well as information from NGOs, websites dedicated in memory to victims, local and state news reports, annual government reports, and sometimes even reports from terrorist organisations themselves. These sources were cross-referenced, and the researchers acknowledge that this could give way to issues of under-reporting. In the dataset, Ideology represents the ideological orientation of the perpetrators, an adaptation of Sánchez-Cuenca and De la Calle (2009), ranging from Vigilante (0), Nationalist (1), Extreme-left (2), Extreme-right (3), and Neonazi (4). Ideology is a codified variable, with each view corresponding to a number 0-4. My outcome variables of interests are ViolenceType and Selectivity. Violencetype represents three forms of political violence: urban political violence (0), diffuse terrorism (1), and organised terrorism (2). This variable is also codified from 0-2. Urban political violence refers to non-planned, often gang-related incidents of violence. Diffuse terrorism refers to smaller-scale organisational killings. Organised terrorism refers to planned attacks carried out by organised groups. Finally, Selectivity represents the degree of selectivity of domestic terrorist attacks. Attacks are classified by Individual (selective) killings (1), State killing (security forces and State officials) (2), group killing (social traits) (3), Society killing (4), Collateral killing (5), or Mistake (6).

My research is cross-sectional and correlational in its design, as it uses data from a particular point in time to draw conclusions. While the data contains information from the years 1965 to 2005, it does not analysis whether domestic terrorists attacks have changed over time, and thus all data points can be treated as a "snapshot." In other words, time was not a variable in the dataset. The dataset focuses on correlation and patterns rather than causal effects due to the nature of the data collection. Thus, there are no explicitly defined independent or dependent variables in the dataset. However, my study seeks whether a causal relationship exists between my independent variable, Ideology, and dependent variables, ViolenceType and Selectivity. If I observe greater numbers of ViolenceType 0 and 1 and Selectivity 1, 2, and 6 in less extreme ideologies (Ideology 0 and 1), this will provide support for my hypothesis. If I observe greater numbers of ViolenceType 1 and 2 and Selectivity 3-5 in more extreme ideologies (Ideology 2, 3, and 4), this will also provide support for my hypothesis. This would be disproved if ViolenceType and Selectivity did not vary across ideology and rather were randomised.

```{r}

#install.packages("distill")
#install.packages("haven")
library(distill)
library(haven)
library(tidyverse)
library(ggplot2)
library(infer)

DTV_1_2009_1 <- read_sav("dataverse_files/DTV 1-2009_1.sav")
DTV_1_2009_2 <- read_sav("dataverse_files/DTV 1-2009_1.sav")
DTV_1_2009_3 <- read_sav("dataverse_files/DTV 1-2009_1.sav")

DTV_1_2009_1 |>
  ggplot(aes(x = violencetype)) +
  geom_bar() +
  labs(title = "Count of Terrorist Organisation Attack Violence Type ", x = "Violence Type", y = "Count")

DTV_1_2009_1 |>
  ggplot(aes(x = selectivity)) +
  geom_bar() +
  labs(title = "Count of Terrorist Organisation Attack Selectivity ", x = "Selectivity", y = "Count")

```

## Results

```{r}

#Graphs

DTV_1_2009_2 |>
  ggplot(aes(x = ideology, y = violencetype))  + labs(title = "Violence Types Across Ideology", x = "Ideology (Vigilante, 0, to Neonazi, 4)", y = "Violence Type") + geom_jitter(alpha = 0.25)

DTV_1_2009_3 |>
  ggplot(aes(x = ideology, y = selectivity))  + labs(title = "Selectivity Across Ideology", x = "Ideology (Vigilante, 0, to Neonazi, 4)", y = "Selectivity") + geom_jitter(alpha = 0.25)

selectivity_regression <- lm(selectivity ~ as.factor(ideology), data = DTV_1_2009_1)
modelsummary::modelsummary(selectivity_regression, statistic = "p.value", 
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))

violencetype_regression <- lm(violencetype ~ as.factor(ideology), data = DTV_1_2009_1)
modelsummary::modelsummary(violencetype_regression, statistic = "p.value", 
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))

```

As seen from the "Violence Types Across Ideology" graph, we see the largest concentration of attacks around Ideology 1 (Nationalist) and Violence Type 2 (organized terrorism). We also see a large concentration of attacks around Ideology 0 (Vigilante) and Violence Type 2 (organised terrorism). When looking across the horizontal axis, it can be said that Violence Type 2 (organised terrorism) is the most concentrated overall compared to Violence Type 0 (urban political violence) and 1 (diffuse terrorism), regardless of ideology. Besides Ideology 4 (neonazi) and Violence Type 1 (diffuse terrorism), there is not a dense concentration of attacks in either Violence Type 0 (urban political violence) or Violence Type 1 (diffuse terrorism). Ideology 0 (Vigilante) and Violence Type 1 (diffuse terrorism) have no points, nor does Ideology 4 (Neonazi) and Violence Type 2 (organised terrorism). When comparing this to the "Selectivity Across Ideology" graph, which measures attack selectivity across ideology, we observe the largest concentration in Ideology 1 (Nationalist) and Selectivity 2 (state killing). The second largest concentration is in Ideology 0 (Vigilante) and Selectivity 3 (group killing). When looking across the horizontal axis, it is less clear what Selectivity is most concentrated, although upon observation it appears to be either Selectivity 2 (state killing) or Selectivity 3 (group killing). 

The first regression model is a regression of selectivity as the dependent variable and ideology as the main independent variable. It uses the as.factor() function for ideology because each codified variable represents a separate entity rather than a continual scope. Ideology 0 (Vigilante) has a coefficient of 2.945, Ideology 1 (Nationalist) is -0.477, Ideology 2 (Extreme left) is -0.544, Ideology 3 (Extreme right) is 0.463, and Ideology 4 (neonazi) is 0.018. In the context of this study, individuals subscribing to Ideology 0, for instance, increased their selectivity of attacks by 2.945 units. The same analysis can be done for individuals of Ideology 1, who decrease their selectivity of attacks by 0.477 units. Coefficients on Ideologies 0, 1, 2, and 3 are statistically significant, whereas Ideology 4 is not. The second regression model is a regression of violence type as the dependent variable and ideology as the main independent variable. Ideology 0 (Vigilante) has a coefficient of 1.994, Ideology 1 (Nationalist) is -0.003, Ideology 2 (Extreme left) is -0.041, Ideology 3 (Extreme right) is -0.115, and Ideology 4 (neonazi) is -1.170. In the context of this study, individuals subscribing to Ideology 2, for instance, decreased their violence type of attacks by -0.041 units. The same analysis can be done for individuals on the other extreme in Ideology 3, who decrease their violence type of attacks by 0.115 units. The most significant decrease in violence type occurs in Ideology 4, as it is the number with the highest magnitude. Coefficients on Ideologies 0, 2, 3, and 4 are statistically significant, whereas Ideology 1 is not. In order to interpret both regression models causally, we would need further analysis, as we would need to control for other variables, replicate the data, and consider alternative explanations and third-variable problems.

## Conclusion

I hypothesised that less extreme ideologies would have less organised and less selective attacks, and more extreme ideologies would have more organised and more selective attacks. There is little overlap between Ideologies 2 (extreme-left), 3 (extreme-right), and 4 (neonazi) and Violence Type 0 (urban political violence). Instead, we see very little victims in this section, instead falling into Violence Type 2 (organised terrorism), which makes sense given that I hypothesised more extreme acts of violence for more extreme ideologies. However, I hypothesised that there would be overlap between Ideology 0/Ideology 1 and ViolenceType 0/ViolenceType 1 because less extreme ideologies would be less violent. In reality, we see practically no overlap between these and instead, Ideology 0/Ideology 1 attacks are concentrated in ViolenceType 2 (organised terrorism), meaning that less extreme ideologies are correlated with more organised types of violence. Overall, a majority of attacks, regardless of ideology, were acts of organised terrorism. In terms of selectivity, I expected an overlap between Ideology 0/Ideology 1 and Selectivity 1, 2, and 6. We do end up seeing a large concentration of Ideology 1 and Selectivity 2, although Ideology 0 is mainly concentrated in Selectivity 3, suggesting that vigilantes are often involved in targeted group killings. We see that more extreme ideologies (2, 3, and 4) tend to be concentrated around wider acts of violence, showcased in the overlap with Selectivity 2, 3, and 4.

A major limitation of my analysis is that correlation does not imply causation. Since this was not a randomised control trial (and rather mimics an observational study,) it is hard to say whether ideology is the direct cause of certain violence types and selectivity. In observational studies, researchers observe a naturally assigned treatment, and it is very commonly the case that you often cannot randomise them for ethical and logistical reasons. This raises questions of internal validity, which addresses if causal assumptions are satisfied and if we can interpret this as a causal effect. RCTs usually have higher internal validity, and observational studies have less because the treatment/control groups differ in ways that are hard to measure. If I could improve my analysis, I would take the time to dissect each individual ideology and compared it to each individual violence type and selectivity to find a regression model. This way, the data and analysis would be more specific and then lay the basis of comparison, as opposed to my current approach, which was to compare all the data before putting it into the regression model. Overall, ideology can be an indicator of the types of violence and the scale at which terrorist organisations conduct their attacks, but it requires more in-depth analysis to prove a definitive causal relationship.
