---
title: "Gov 50 Final Project"
author: "Emma Lucas"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Introduction

For my final project, I was interested in exploring data related to domestic terrorism, violence, and ideology. How does the ideological orientation of domestic terrorist perpetrators influence, if at all, the type of political violence and the selectivity of said violence? In this study, I examined whether or not certain types and selectivity of terrorist violence can indicate patterns and trends within specific ideologies. 

I hypothesise that less extreme ideologies will have less organised and wide-scale types of violence and that they will be less selective in their attacks. Less extreme ideologies, while still passionate about their causes, are less likely to take greater steps to ensure their causes. The more the ideology deviates from the norm, the more inclined terrorist organisations are to act out to bring attention to their cause. I also predict that less extreme ideologies will be smaller-scaled in terms of selectivity whereas more extreme ideologies will be larger-scaled. My explanatory variable of interest is what ideology the terrorist group aligns with. 

## Data Section

The dataset currently holds information on all terrorist attacks from 1965 to 2005 carried out by domestic terrorist organizations in European nations. Each data point represents an individually killed as a directly result of the terrorist attack. To gather the data, researchers relied on pre-existing datasets of terrorist killings (mainly TWEED and Global Terrorism Database), as well as information from NGOs, websites dedicated in memory to victims, local and state news reports, annual government reports, and sometimes even reports from terrorist organisations themselves. These sources were cross-referenced, and the researchers acknowledge that this could give way to issues of under-reporting.

My research is cross-sectional, as it uses data from a particular point in time to draw conclusions. While the data contains information from the years 1965 to 2005, it does not analysis whether domestic terrorists attacks have changed over time, and thus all data points can be treated as a "snapshot." In other words, time was not a variable in the dataset. The dataset focuses on correlation and patterns rather than causal effects due to the nature of the data collection. Thus, there are no explicitly defined independent or dependent variables in the dataset. However, my study tests whether a causal relationship could exist between the independent variable, Ideology, and dependent variables, ViolenceType and Selectivity.

In the dataset, Ideology represents the ideological orientation of the perpetrators, an adaptation of Sánchez-Cuenca and De la Calle (2009), ranging from Vigilante (0), Nationalist (1), Extreme-left (2), Extreme-right (3), and Neonazi (4). Ideology is a codified variable, with each view corresponding to a number 0-4. My outcome variables of interests are ViolenceType and Selectivity. Violencetype represents three forms of political violence: urban political violence (0), diffuse terrorism (1), and organised terrorism (2). This variable is also codified from 0-2. Urban political violence refers to non-planned, often gang-related incidents of violence. Diffuse terrorism refers to smaller-scale organisational killings. Organised terrorism refers to planned attacks carried out by organised groups. Finally, Selectivity represents the degree of selectivity of domestic terrorist attacks. Attacks are classified by Individual (selective) killings (1), State killing (security forces and State officials) (2), group killing (social traits) (3), Society killing (4), Collateral killing (5), or Mistake (6). 

If I observe greater numbers of ViolenceType 0-1 and Selectivity 1-3 and 6 in less extreme ideologies, this will provide support for my hypothesis. If I observe greater numbers of ViolenceType 1-2 and Selectivity 3-5 in more extreme ideologies, this will also provide support for my hypothesis. This would be disproved if ViolenceType and Selectivity did not vary across ideology and rather were randomised.

```{r}

#install.packages("distill")
#install.packages("haven")
library(distill)
library(haven)
library(tidyverse)

DTV_1_2009_1 <- read_sav("dataverse_files/DTV 1-2009_1.sav")

DTV_1_2009_2 <- read_sav("dataverse_files/DTV 1-2009_1.sav")

DTV_1_2009_3 <- read_sav("dataverse_files/DTV 1-2009_1.sav")

library(ggplot2)

DTV_1_2009_1 |>
  ggplot(aes(x = ideology)) +
  geom_bar() +
  labs(title = "Count of Terrorist Group's Ideology ", x = "Ideology (Vigilante, 0, to Neonazi, 4)", y = "Count")

DTV_1_2009_2 |>
  ggplot(aes(x = ideology, y = violencetype))  + labs(title = "Violence Types Across Ideology", x = "Ideology (Vigilante, 0, to Neonazi, 4)", y = "Violence Type") + geom_jitter(alpha = 0.25)

DTV_1_2009_3 |>
  ggplot(aes(x = ideology, y = selectivity))  + labs(title = "Selectivity Across Ideology", x = "Ideology (Vigilante, 0, to Neonazi, 4)", y = "Selectivity") + geom_jitter(alpha = 0.25)

```

```{r}

ggplot(DTV_1_2009_1, aes(x = ideology, y = violencetype)) +
  geom_jitter(alpha = 0.25) +
  labs(
    x = "Ideology",
    y = "Violence Type",
    title = "Ideology and Violence Type"
  ) +
  stat_summary_bin(fun = "mean", color = "indianred1", size = 3,
                   geom = "point", binwidth = 1)

```


```{r}

regression <- lm(ideology ~ selectivity + violencetype,
               data = DTV_1_2009_1)

modelsummary::modelsummary(regression,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))

```


# Analysis

As seen from the first graph, there are almost triple the number of terrorist groups with Ideology 1 (Nationalist(pro-secession)) compared to the second largest pool of terrorist groups, Ideology 0 (Vigilante (against secession)). From greatest to least with the remaining groups, Ideology 3 (Extreme-right), Ideology 2 (Extreme-left), and Ideology 4 (Neonazi) are a bit less than half that of ideology 0.

In the second graph, we see the largest concentration of attacks around Ideology 1 and Violence Type 2 (Organized Terrorism). We also see a large concentration of attacks around Ideology 0 and Violence Type 2. When looking across the horizontal axis, it can be said that Violence Type 2 is the most concentrated overall, regardless of ideology. Besides Ideology 4 and Violence Type 1 (Diffuse terrorism), there is not a dense concentration of attacks. Ideology 0 and Violence Type 1 have no points, nor does Ideology 4 and Violence Type 2.

When comparing this to the third graph, which measures attack selectivity across ideology, we observe the largest concentration in Ideology 1 and Selectivity 2 (State killing (security forces and State officials)). The second largest concentration is in Ideology 0 and Selectivity 3 (Social group killing (social traits)). When looking across the horizontal axis, it is less clear what Selectivity is most concentrated, although upon observation it appears to be either Selectivity 2 or Selectivity 3. 
